### 2021-06-01 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Edits to fig layouts mainly

\chapter{Disk slices and trapeziums}

@@

\begin{figure}[tp]
	\centering % <-- added
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[scale=0.5]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
		\centering
		\caption{The Laplacian operator $\laplacewiii$}
	\end{subfigure}\hfil % <-- added
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[scale=0.5]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
		\centering
		\caption{A variable coefficient Helmholtz operator}
	\end{subfigure}\hfil % <-- added

	\medskip
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[scale=0.5]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
		\centering
		\caption{The biharmonic operator $\biharmonictwo$}
	\end{subfigure}\hfil % <-- added
    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity. For (b), the weighted variable coefficient Helmholtz operator is $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.}
	\label{fig:sparsity}
\end{figure}

@@

In this chapter we have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve -- notably here the disk-slice -- with Dirichlet boundary conditions. This work extends the triangle case \cite{beuchler2006new,li2010optimal,olver2019triangle} and whole disk case \cite{vasil2016tensor} \bstodo{cite more here} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements (for example, by dividing the disk or a section of the disk into disk-slice elements, which has applications in turbulent pipe flow \cite{eggels1994fully, vasil2016tensor}).

>>>>

\chapter{Disk slices and trapeziums}\label{CHAPTER:diskslice}

@@

\begin{figure}[tp]
	\centering % <-- added
	\begin{subfigure}{0.5\textwidth}
		\centerline{\includegraphics[scale=0.5]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}}
		\centering
		\caption{}
	\end{subfigure}\hfil % <-- added
	\begin{subfigure}{0.5\textwidth}
		\centerline{\includegraphics[scale=0.5]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}}
		\centering
		\caption{}
	\end{subfigure}\hfil % <-- added

	\medskip
	\begin{subfigure}{0.5\textwidth}
		\centerline{\includegraphics[scale=0.5]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}}
		\centering
		\caption{}
	\end{subfigure}\hfil % <-- added
    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity. (a) The Laplacian operator $\laplacewiii$. (b) The variable coefficient Helmholtz operator $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$. (c) The biharmonic operator $\biharmonictwo$}
	\label{fig:sparsity}
\end{figure}

@@

In this chapter we have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve -- notably here the disk-slice -- with Dirichlet boundary conditions. This work extends the triangle case \cite{beuchler2006new, li2010optimal, olver2019triangle} and whole disk case \cite{boyd2011comparing, vasil2016tensor} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements (for example, by dividing the disk or a section of the disk into disk-slice elements, which has applications in turbulent pipe flow \cite{eggels1994fully, vasil2016tensor}).

### 2021-06-01 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Fixed appendix numbering
Edits to fig layouts mainly

\chapter{Spherical caps}

@@

We consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): in spherical coordinates 
\bseqn{
	z &= \cosphi, \\
	x &= \sinphi \costheta = \rho(z) \costheta, \\
	y &= \sinphi \sintheta = \rho(z) \sintheta.
}
where $ \rho(z) := \sqrt{1-z^2}$, we have
\bseqn{
	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo
}
i.e. $\DeltaS f(\xvec) = \Delta f({\xvec \over \norm{\xvec}})$ for $\xvec := (x,y,z) \in \R^3$. We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see \bsreftheorem{theorem:sc:sparsityofdifferentialoperators}. Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle}, disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations involving Jacobi polynomials on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}. Just as we proceeded in \bsrefchapter{chapter:diskslice} for the disk-slice case in 2D (see also \cite{snowball2019sparse}), we will use an integration-by-parts argument to deduce the sparsity structure.

The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical \cite[\S5]{magnus1995painleve}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals have already been outlined in \bsrefchapter{chapter:diskslice} (see also \cite{snowball2019sparse}). In particular, by exploiting the connection with these 1D OPs we can construct discretizations of general partial differential operators of size $(p+1)^2 \times (p+1)^2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree. This clearly compares favourably to proceeding in a na\"ive approach where one would require $O(p^6)$ operations.

@@

By construction, $\{\scopnkia\}$ are orthogonal with respect to the inner product
\bseqn{
	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D A \\
	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
}
with
\bseqnnumber{
	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} = \pi \: \normgenjac^{(a, 2k)}. \label{eqn:sc:normscop}
}
A method for obtaining explicit recurrence coefficients and evaluating integrals for the weight function $\genjacw^{(a,b)}(x)$ was established in \bsrefchapter{chapter:diskslice} (see also \cite{snowball2019sparse}). The weight is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}.

@@

Recall that we can express the three-term recurrences associated with $\genjac_n^{(a,b)}$ as
\bseqnnumber{
	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:sc:Rrecurrence}
}
where the coefficients are calculable (see \bsrefchapter{chapter:diskslice}). We can use \bsrefeqn{eqn:sc:Rrecurrence} to determine the 3D recurrences for $\scopnkia(x,y,z)$. Importantly, we can deduce sparsity in the recurrence relationships.  We first require the following lemma.

@@

\begin{figure}[tp]
	\centering % <-- added
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[scale=0.5]{sparsity-of-laplacian}
		\centering
		\caption{The Laplace-Beltrami operator $\Delta^{(1)}_W$}
	\end{subfigure}\hfil % <-- added
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[scale=0.5]{sparsity-of-rho2laplacian}
		\centering
		\caption{The $\rho^2$-factored Laplace-Beltrami operator $D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2$}
	\end{subfigure}\hfil % <-- added

	\medskip
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[scale=0.5]{sparsity-of-helmholtz}
		\centering
		\caption{A variable coefficient Helmholtz operator}
	\end{subfigure}\hfil % <-- added
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[scale=0.5]{sparsity-of-biharmonic}
		\centering
		\caption{The Biharmonic operator $\calB_W^{(2)}$} 
	\end{subfigure}\hfil % <-- added
	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity. For (c), the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.}
	\label{fig:sc:sparsity}
\end{figure}

@@

\begin{figure}[tp]
	\centering
	\begin{subfigure}[b]{0.55\textwidth}
		\includegraphics[scale=0.4]{condition-numbers-laplacian-log-N=200}
	\end{subfigure}
	\hfill%

	\begin{subfigure}[b]{0.55\textwidth}
		\includegraphics[scale=0.4]{condition-numbers-max-laplacian-log}
	\end{subfigure}
	\hfill%
	\caption{Plots of the condition number for the Laplacian operator matrix $\Delta^{(1)}_W$ and the preconditioned matrix $P^{-1} \Delta^{(1)}_W$ where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$. Top: The condition numbers of each diagonal block of $\Delta^{(1)}_W$ for $N = 200$, as well as those for $P^{-1} \Delta^{(1)}_W$. Bottom: The maximum condition number of all diagonal blocks of the Laplacian and the preconditioned Laplacian operators, as the degree $N$ increases.}
	\label{fig:sc:conditionnumbers}
\end{figure}

@@

Firstly, we note that
\bseqn{
	\int_0^{2\pi} g\big(\cos \theta, \sin\theta\big) \: \D \theta &= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}} \\
	&= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(-x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}}
}
for any function $g$, using a change of variables $x \to -x$ for the second term in the integral. Then, integrating the even function $f_e$ we have
\bseqn{
	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \\
	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z\big) \: \D \theta \Big) \: \D z \\
	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + f_e\big(-\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \Big] \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
	&\quad \quad \approx \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) s_l, \rho(z) \sqrt{1 - s_l^2}, z\big) \Big) \: \D z \quad (\star) \\
	&\quad \quad \approx \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) s_l, \rho(t_j) \sqrt{1 - s_l^2}, t_j\big) \quad (\star \star) \\
	&\quad \quad = \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
}
Suppose $f$ is a polynomial in $x,y,z$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial. It follows that $s \mapsto f_e\big(\rho(z)s,\rho(z)\sqrt{1-s^2}, z\big)$ for fixed $z$ is then a polynomial of degree $\le N$. We therefore achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.

@@

Fix $a \in \R$. Then for any function $f : \Omega \to \R$ we can express $f$ by
\bseqn{
	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \bm{f}_k = \bigscopNa(x,y,z)^\top \: \bm{f}
}
for N sufficiently large, where $\bigscopNka, \bigscopNa$ is defined in \bsrefeqns{eqn:sc:OPdefNka}{eqn:sc:OPdefNa} and where

\bseqn{
	\bm{f}_k &:= 
		\begin{pmatrix}
			f_{k,k,0} \\
			f_{k,k,1} \\
			\vdots \\
			f_{N,k,0} \\
			f_{N,k,1}
		\end{pmatrix} \in \R^{2(N-k+1)} \quad \text{for } n = 1,2,\dots,N, \quad
	\bm{f}_0 := 
		\begin{pmatrix}
			f_{0,0,0} \\
			\vdots \\
			f_{N,0,0}
		\end{pmatrix} \in \R^{N+1}, \\
	\bm{f} &:= 
		\begin{pmatrix}
			\bm{f}_0 \\
			\vdots \\
			\bm{f}_N
		\end{pmatrix} \in \R^{2(N+1)^2}, \quad \quad
	f_{n,k,i} := \ip< f, \: \scopnkia>_{\scopa} \: \norm{\scopnkia}^{-2}_{\scopa}.
}
Recall from \bsrefeqn{eqn:sc:normscop} that $\norm{\scopnkia}^2_{\scopa} = \normgenjac^{(a,2k)} \: \pi$. Using the quadrature rule detailed in \bsrefsection{section:sc:quadrule} for the inner product, we can calculate the coefficients $f_{n,k,i}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$, $i = 0,1$: 
\bseqn{
	f_{n,k,i} &= \frac{1}{2 \: \normgenjac^{(a,2k)} \: \pi} \sum_{j=1}^{M} w_j \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) +f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big] \\
	&= \frac{1}{M_2 \: \normgenjac^{(a,2k)}} \sum_{j=1}^{M} \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) +f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big]
}
where the quadrature nodes and weights are those from \bsreftheorem{theorem:sc:quadrule}, and $M = M_1 M_2$ with $2M_1 - 1 \ge N, M_2 - 1 \ge N$ (i.e. we can choose $M_2 := N + 1$ and $M_1 := \ceil{\frac{N+1}{2}}$).

@@

\begin{figure}[tp]
	\centering % <-- added
	\includegraphics[scale=0.4]{complexity-Nend=1000}
	\caption{Time in seconds to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$, for a rotationally invariant $v(x,y,z) = v(z)$. This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution (the number of unknowns is then $(N+1)^2$). Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$ and $v(x,y,z) = v(z) = \cos(z)$.}
	\label{fig:sc:complexity}
\end{figure}

>>>>

\chapter{Spherical caps}\label{CHAPTER:sphericalcaps}

@@

We consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): in spherical coordinates 
\bseqn{
	z &= \cosphi, \\
	x &= \sinphi \costheta = \rho(z) \costheta, \\
	y &= \sinphi \sintheta = \rho(z) \sintheta.
}
where $ \rho(z) := \sqrt{1-z^2}$, we have
\bseqn{
	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo
}
i.e. $\DeltaS f(\xvec) = \Delta f({\xvec \over \norm{\xvec}})$ for $\xvec := (x,y,z) \in \R^3$. We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see \bsreftheorem{theorem:sc:sparsityofdifferentialoperators}. Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle}, disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations involving Jacobi polynomials on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}. Just as we proceeded in \bsrefchapter{CHAPTER:diskslice} for the disk-slice case in 2D (see also \cite{snowball2019sparse}), we will use an integration-by-parts argument to deduce the sparsity structure.

The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical \cite[\S5]{magnus1995painleve}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals have already been outlined in \bsrefchapter{CHAPTER:diskslice} (see also \cite{snowball2019sparse}). In particular, by exploiting the connection with these 1D OPs we can construct discretizations of general partial differential operators of size $(p+1)^2 \times (p+1)^2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree. This clearly compares favourably to proceeding in a na\"ive approach where one would require $O(p^6)$ operations.

@@

By construction, $\{\scopnkia\}$ are orthogonal with respect to the inner product
\bseqn{
	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D A \\
	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
}
with
\bseqnnumber{
	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} = \pi \: \normgenjac^{(a, 2k)}. \label{eqn:sc:normscop}
}
A method for obtaining explicit recurrence coefficients and evaluating integrals for the weight function $\genjacw^{(a,b)}(x)$ was established in \bsrefchapter{CHAPTER:diskslice} (see also \cite{snowball2019sparse}). The weight is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}.

@@

Recall that we can express the three-term recurrences associated with $\genjac_n^{(a,b)}$ as
\bseqnnumber{
	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:sc:Rrecurrence}
}
where the coefficients are calculable (see \bsrefchapter{CHAPTER:diskslice}). We can use \bsrefeqn{eqn:sc:Rrecurrence} to determine the 3D recurrences for $\scopnkia(x,y,z)$. Importantly, we can deduce sparsity in the recurrence relationships.  We first require the following lemma.

@@

\begin{figure}[tp]
	\centering % <-- added
	\begin{subfigure}{0.5\textwidth}
		\centerline{\includegraphics[scale=0.5]{sparsity-of-laplacian}}
		\centering
		\caption{The Laplace-Beltrami operator $\Delta^{(1)}_W$}
	\end{subfigure}\hfil % <-- added
	\begin{subfigure}{0.5\textwidth}
		\centerline{\includegraphics[scale=0.5]{sparsity-of-rho2laplacian}}
		\centering
		\caption{The $\rho^2$-factored Laplace-Beltrami operator $D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2$}
	\end{subfigure}\hfil % <-- added

	\medskip
	\begin{subfigure}{0.5\textwidth}
		\centerline{\includegraphics[scale=0.5]{sparsity-of-helmholtz}}
		\centering
		\caption{A variable coefficient Helmholtz operator}
	\end{subfigure}\hfil % <-- added
	\begin{subfigure}{0.5\textwidth}
		\centerline{\includegraphics[scale=0.5]{sparsity-of-biharmonic}}
		\centering
		\caption{The Biharmonic operator $\calB_W^{(2)}$} 
	\end{subfigure}\hfil % <-- added
	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity. For (c), the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.}
	\label{fig:sc:sparsity}
\end{figure}

@@

\begin{figure}[tp]
	\centering
	\begin{subfigure}[b]{0.55\textwidth}
		\centerline{\includegraphics[scale=0.4]{condition-numbers-laplacian-log-N=200}}
	\end{subfigure}
	\hfill%

	\begin{subfigure}[b]{0.55\textwidth}
		\centerline{\includegraphics[scale=0.4]{condition-numbers-max-laplacian-log}}
	\end{subfigure}
	\hfill%
	\caption{Plots of the condition number for the Laplacian operator matrix $\Delta^{(1)}_W$ and the preconditioned matrix $P^{-1} \Delta^{(1)}_W$ where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$. Top: The condition numbers of each diagonal block of $\Delta^{(1)}_W$ for $N = 200$, as well as those for $P^{-1} \Delta^{(1)}_W$. Bottom: The maximum condition number of all diagonal blocks of the Laplacian and the preconditioned Laplacian operators, as the degree $N$ increases.}
	\label{fig:sc:conditionnumbers}
\end{figure}

@@

Firstly, we note that
\bseqn{
	\int_0^{2\pi} g\big(\cos \theta, \sin\theta\big) \: \D \theta &= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}} \\
	&= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(-x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}}
}
for any function $g$, using a change of variables $x \to -x$ for the second term in the integral. Then, integrating the even function $f_e$ we have
\bseqn{
	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \\
	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z\big) \: \D \theta \Big) \: \D z \\
	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + f_e\big(-\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \Big] \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
	&\quad \quad \approx 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) s_l, \rho(z) \sqrt{1 - s_l^2}, z\big) \Big) \: \D z \quad (\star) \\
	&\quad \quad \approx 2 \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) s_l, \rho(t_j) \sqrt{1 - s_l^2}, t_j\big) \quad (\star \star) \\
	&\quad \quad = 2 \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
}
Suppose $f$ is a polynomial in $x,y,z$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial. It follows that $s \mapsto f_e\big(\rho(z)s,\rho(z)\sqrt{1-s^2}, z\big)$ for fixed $z$ is then a polynomial of degree $\le N$. We therefore achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.

@@

Fix $a \in \R$. Then for any function $f : \Omega \to \R$ we can express $f$ by
\bseqn{
	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \bm{f}_k = \bigscopNa(x,y,z)^\top \: \bm{f}
}
for N sufficiently large, where $\bigscopNka, \bigscopNa$ is defined in \bsrefeqns{eqn:sc:OPdefNka}{eqn:sc:OPdefNa} and where
\bseqn{
	\bm{f}_k &:= 
		\begin{pmatrix}
			f_{k,k,0} \\
			f_{k,k,1} \\
			\vdots \\
			f_{N,k,0} \\
			f_{N,k,1}
		\end{pmatrix} \in \R^{2(N-k+1)} \quad \text{for } n = 1,2,\dots,N, \quad
	\bm{f}_0 := 
		\begin{pmatrix}
			f_{0,0,0} \\
			\vdots \\
			f_{N,0,0}
		\end{pmatrix} \in \R^{N+1}, \\
	\bm{f} &:= 
		\begin{pmatrix}
			\bm{f}_0 \\
			\vdots \\
			\bm{f}_N
		\end{pmatrix} \in \R^{2(N+1)^2}, \quad \quad
	f_{n,k,i} := \ip< f, \: \scopnkia>_{\scopa} \: \norm{\scopnkia}^{-2}_{\scopa}.
}
Recall from \bsrefeqn{eqn:sc:normscop} that $\norm{\scopnkia}^2_{\scopa} = \normgenjac^{(a,2k)} \: \pi$. Using the quadrature rule detailed in \bsrefsection{section:sc:quadrule} for the inner product, we can calculate the coefficients $f_{n,k,i}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$, $i = 0,1$: 
\bseqn{
	f_{n,k,i} &= \frac{1}{\normgenjac^{(a,2k)} \: \pi} \sum_{j=1}^{M} w_j \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) \\
	&\quadeight \quad + f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big] \\
	&= \frac{1}{M_2 \: \normgenjac^{(a,2k)}} \sum_{j=1}^{M_1} w^{(t)}_j \sum_{l=1}^{M_2} \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) \\
	&\quadeight \quadfour + f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big]
}
where the quadrature nodes and weights are those from \bsreftheorem{theorem:sc:quadrule}, and $M = M_1 M_2$ with $2M_1 - 1 \ge N, M_2 - 1 \ge N$ (i.e. we can choose $M_2 := N + 1$ and $M_1 := \ceil{\frac{N+1}{2}}$).

@@

\begin{figure}[tp]
	\centering % <-- added
	\centerline{\includegraphics[scale=0.4]{complexity-Nend=1000}}
	\caption{Time in seconds to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$, for a rotationally invariant $v(x,y,z) = v(z)$. This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution (the number of unknowns is then $(N+1)^2$). Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$ and $v(x,y,z) = v(z) = \cos(z)$.}
	\label{fig:sc:complexity}
\end{figure}
